---
title: "United States Land Use Analysis 1945 - 2012"
purpose: ECON6010 Final Project -- Fall Semester 2019
author: "Chris Bucher, Xiaoxiang Chen, Albert Gustafson, Brady Kraemer, and Sam Ransohoff"
date: "11/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Downloads")
```

'''{r setup, include = FALSE}


<br />

# Introduction

### Research Question
* How has land use in the United States changed over time? 
  + Technological advances
  + Labor force inconsistencies
  + Growing populations
    + Increased food demand

### What happened to all the farmers?

* Between 1948 and 2015, total farm output has more than doubled in the United States. (USDA) 
* What about the productivity of land?

### What happened happened to all the land?

* If there’s less cropland in the US now than there was 70 years ago, what has taken its place?
* Environmental Kuznets curve?
* Urbanization during the same period means that more people were living and working in cities
* Which effect dominates?
  

# Packages Used
* dplyr      -->  Data manipulation grammar
* tidyverse  -->  Provides core packages to use in data analysis
* tidyr      -->  Tidies messy data
* reshape    -->  Transforms data from short and wide to long and thin format or vice-versa
* usmap      -->  U.S. map data frames of varying region types for mapping purposes
* ggplot2    -->  Create graphics
* plotly     -->  Create interactive graphs
* gridExtra  -->  Arranges plots into grid format

```{r Packages, include = FALSE}

library(dplyr)    
library(tidyverse)  
library(tidyr)      
library(reshape)    
library(usmap)      
library(ggplot2)    
library(plotly)     
library(gridExtra)
library(readxl)
```

# Data Sources

* Land Use Data was obtained from the United States Department of Agriculture (USDA)
* Employment data was obtained from Federal Reserve Economic Data (FRED)

```{r Data Cleaning Pt. 1, include = FALSE}

########
## LAND USE DATA
########

# create a function that negates %in% so that we can exclude records that don't match
`%nin%` = Negate(`%in%`)

# Import land use data from USDA file
all_land_use <- read.csv('https://www.ers.usda.gov/webdocs/DataFiles/52096/MajorLandUse.csv?v=8645.3', stringsAsFactors = FALSE, header = TRUE) 

# Get rid of Alaska and Hawaii since they have NAs:
drop_states <- c("Alaska", "Hawaii")

all_land_use <- all_land_use %>% filter(Region.or.State %nin% drop_states)

# Remove unnecessary columns
all_land_use <- all_land_use %>% 
  dplyr::select(-c("SortOrder", "Region","Cropland.used.for.crops", "Cropland.idled", "Cropland.used.for.pasture", 
                   "Forest.use.land.grazed", "Forest.use.land.not.grazed", "Land.in.rural.transportation.facilities",
                   "Land.in.defense.and.industrial.areas", "Farmsteads..roads..and.miscellaneous.farmland"))
  

# Rename columns
colnames(all_land_use) <- c('state', 'year', 'total_land', 'total_cropland', 'total_grassland', 'total_forest', 'total_special', 'parks_and_wildlife', 'total_urban_use', 'total_other')

# Convert column contents to numeric values:
all_land_use$total_land <- as.numeric(all_land_use$total_land)
all_land_use$total_cropland <- as.numeric(all_land_use$total_cropland)
all_land_use$total_forest <- as.numeric(all_land_use$total_forest)
all_land_use$total_special <- as.numeric(all_land_use$total_special)
all_land_use$parks_and_wildlife <- as.numeric(all_land_use$parks_and_wildlife)
all_land_use$total_urban_use <- as.numeric(all_land_use$total_urban_use)
all_land_use$total_other <- as.numeric(all_land_use$total_other)
all_land_use$total_grassland <- as.numeric(all_land_use$total_grassland)

#Sum cropland and grassland

all_land_use$total_cropland = all_land_use$total_cropland + all_land_use$total_grassland
all_land_use <- all_land_use %>% select(-c("total_grassland"))

########
# CREATE PERCENTAGE DATA FRAME
########

# Create DF that is percentage data:
pct_all_land_use <- all_land_use %>% mutate(cropland_pct = (total_cropland/total_land)*100,
                                            forest_pct = (total_forest/total_land)*100, 
                                            special_pct = (total_special/total_land)*100, 
                                            parks_wild_pct = (parks_and_wildlife/total_land)*100,
                                            urban_pct = (total_urban_use/total_land)*100, 
                                            other_pct = (total_other/total_land)*100)

# Remove non percentages from pct_all_land_use DF:
pct_all_land_use <- pct_all_land_use %>% select(-c('total_land', 'total_cropland', 'total_forest', 'total_special',
                                                   'parks_and_wildlife', 'total_urban_use', 'total_other'))

########
# CREATE DF FOR REGIONS, STATES, US TOTAL:
########

# Create pct and acreage DF of all US total data:
us_total_pct <- pct_all_land_use %>% subset(state == "U.S. total")  # US total data in pct
us_total_acre <- all_land_use %>% subset(state == 'U.S. total')     # US total data in acres

# Create pct and acreage DF of states:
state_names <- c(state.name) #list created of state names in the USA
state_pct <- subset(pct_all_land_use, state %in% state_names)       # State data in pct
state_acre <- subset(all_land_use, state %in% state_names)          # State data in acres
    # NOTE: This eliminates the 'lower 48' and Washington DC from all DFs ***

# Create pct and acreage DF of regions:
region_names <- c("Northeast", "Lake States", "Corn Belt", "Northern Plains", "Appalachian", "Southeast", "Delta States", 
                  "Southern Plains", "Mountain", "Pacific")
region_pct <- subset(pct_all_land_use, state %in% region_names)     # Region data in pct
names(region_pct)[names(region_pct)== 'state'] <- 'region'          # Rename 'state' -> 'region'
region_acre <- subset(all_land_use, state %in% region_names)        # Region data in acre
names(region_acre)[names(region_acre)== 'state'] <- 'region'          # Rename 'state' -> 'region'


#################
#EMPLOYMENT DATA#
#################

PIE_DATA <- read_excel("PIE_DATA.xlsx")
MajorLandUse <- read.csv("MajorLandUse.csv", stringsAsFactors = FALSE, header = TRUE)

#And here we combine land and employment data

Total_land_use <- MajorLandUse%>%
  filter(SortOrder>=934&SortOrder<=945)%>%
  select(-(1:3))

combined_data <- Total_land_use%>%
  left_join(PIE_DATA, by=c("Year"="YEAR"))

combined_data$Total.cropland <- as.numeric(combined_data$Total.cropland)
combined_data$Total.land <- as.numeric(combined_data$Total.land)

labor_land_comb_data <- combined_data %>% mutate(ag_land_pct = (Total.cropland/Total.land)*100,
                                                 ag_emp_pct = (Ag_workers/Labor_Force)*100) %>% 
  select(Year, ag_land_pct, ag_emp_pct)

```



---

---

---
---


# Land Use Analysis and Visualization

## U.S. Land Use Over Time 

```{r New Variable Creation, echo= FALSE}
#Create a new variable to make the line graphs that will plot land use over time
line_graph_data <- all_land_use %>% select(state, year, total_land, total_cropland, 
total_forest, total_forest, total_special,  parks_and_wildlife, total_urban_use, total_other) %>% filter(state =="U.S. total")
```


```{r Bar Graph of PCT Land Use 1945 vs 2012, echo=FALSE}

# Define names for columns for the bar chart:
bar_chart_titles <- c("State", "Year", "Cropland", "Forest", "Special Use", "Parks & Wildlife", "Urban", "Other")

# Make df of US land use pct in 1945:
Luse1945 <- subset(us_total_pct, year == 1945) # 
  colnames(Luse1945) <- bar_chart_titles          # renames column names to better names for visualization
  Luse1945 <- reshape::melt(Luse1945, id = c('State', 'Year'))  # transposes df shape
  names(Luse1945)[names(Luse1945)== "value"] <- "use_1945"    # renames "value" --> use_1945
  
# Make df of US land use pct in 2012:  
Luse2012 <- subset(us_total_pct, year == 2012)
  colnames(Luse2012) <- bar_chart_titles          # renames column names to better names for visualization
  Luse2012 <- reshape::melt(Luse2012, id = c('State', 'Year'))  # transpose df shape
  names(Luse2012)[names(Luse2012)== "value"] <- "use_2012"    # renames "value" --> use_2012
  
  
# Combine Luse1945 and Luse2012 into single DF:
chris_bar_data <- merge(Luse1945, Luse2012, by = "variable") %>% select(c("variable", "use_1945", 'use_2012'))


# Create the graph
land_use_pct_1945_vs_2012 <- plot_ly(chris_bar_data, x = ~variable, y = ~use_1945, type = 'bar', name = '1945') %>%
  add_trace(y = ~use_2012, name = '2012') %>%
  layout(yaxis = list(title = 'Pecentage'), xaxis = list(title = ''),barmode = 'group')

land_use_pct_1945_vs_2012

```

```{r US Land Use Over Time,echo= FALSE}
#plot all land uses vs. year
  ggplot(line_graph_data, aes(x = year)) + 
  geom_line(aes(y = total_cropland), color = "greenyellow") + 
  geom_line(aes(y = total_forest), color = "darkgreen") +
  geom_line(aes(y = total_special), color = "purple2") +
  geom_line(aes(y = parks_and_wildlife), color = "deepskyblue1") + 
  geom_line(aes(y = total_urban_use), color = "grey42") +
  geom_line(aes(y = total_other), color = "orange") + 
  xlab("Year") + ylab("Land in Thousand Acres") +
  # ggtitle("US Land Use Over Time")    # dont ned the title as it's already above the plotly chart
  scale_x_continuous(breaks=c(1945, 1954, 1964, 1974, 1982, 1992, 2002, 2012))
```


```{r Cropland PCT DFs, message=FALSE, include= FALSE}

pct_cropland <- state_pct  %>% select(state, year, cropland_pct) %>% arrange(year) #pct cropland, state, and year
crop_1945_state <- state_pct %>% select(c("state", "year", "cropland_pct")) %>% subset(year == 1945) 
crop_1964_state <- state_pct %>% select(c("state", "year", "cropland_pct")) %>% subset(year == 1964) 
crop_1992_state <- state_pct %>% select(c("state", "year", "cropland_pct")) %>% subset(year == 1992)
crop_2012_state <- state_pct %>% select(c("state", "year", "cropland_pct")) %>% subset(year == 2012) 

```

<br />  


## Cropland 

* Cropland is a combination of cropland and grassland as defined by the USDA, which are both used for agricultural use
<br />

```{r Cropland Over Time, echo = FALSE}
#plot cropland vs. year
ggplot(line_graph_data, aes(x = year, y = total_cropland)) + geom_line(color = "greenyellow") + geom_point() + ggtitle("Cropland Over Time") + xlab("Year") + ylab("Cropland in Thousand Acres") + scale_x_continuous(breaks=c(1945, 1954, 1964, 1974, 1982, 1992, 2002, 2012))
```

<br /> 

```{r Cropland Maps, echo =FALSE,}

# 1945 Cropland PCT Map:
crop_map_1945 <- plot_usmap(exclude = c("HI", "AK"), data = crop_1945_state, values = "cropland_pct", color = 'black') +
  scale_fill_continuous(low = 'white', high = 'green', name = 'Cropland %', label = scales::comma) +
  theme(legend.position = 'right') +  labs(title = "1945")

# 1964 Cropland PCT Map:
crop_map_1964 <- plot_usmap(exclude = c("HI", "AK"), data = crop_1964_state, values = "cropland_pct", color = 'black') +
  scale_fill_continuous(low = 'white', high = 'green', name = 'Cropland %', label = scales::comma) +
  theme(legend.position = 'right') +  labs(title = "1964")

# 1992 Cropland PCT Map:
crop_map_1992 <- plot_usmap(exclude = c("HI", "AK"), data = crop_1992_state, values = "cropland_pct", color = 'black') +
  scale_fill_continuous(low = 'white', high = 'green', name = 'Cropland %', label = scales::comma) +
  theme(legend.position = 'right') +  labs(title = "1992")

# 2012 Cropland PCT Map:
crop_map_2012 <- plot_usmap(exclude = c("HI", "AK"), data = crop_2012_state, values = "cropland_pct", color = 'black') +
  scale_fill_continuous(low = 'white', high = 'green', name = 'Cropland %', label = scales::comma) +
  theme(legend.position = 'right') +  labs(title = "2012")

grid.arrange(crop_map_1945, crop_map_1964, crop_map_1992, crop_map_2012, ncol=2, nrow=2, top= "U.S. Percentage Cropland 1945-2012")

```

<br /> 

## Woodland


```{r Woodland PCT DFs, message=FALSE, include= FALSE}

## Combine forest_pct and parks_wild_pct to get total woodland percentage (woodlands_pct) in states:
woodland_pct <- state_pct %>% mutate(woodlands_pct = forest_pct + parks_wild_pct) %>% 
  select(state, year, woodlands_pct) %>% arrange(year)

wood_1945_state <- woodland_pct %>% select(c("state", "year", "woodlands_pct")) %>% subset(year == 1945) 
wood_1964_state <- woodland_pct %>% select(c("state", "year", "woodlands_pct")) %>% subset(year == 1964) 
wood_1992_state <- woodland_pct %>% select(c("state", "year", "woodlands_pct")) %>% subset(year == 1992)
wood_2012_state <- woodland_pct %>% select(c("state", "year", "woodlands_pct")) %>% subset(year == 2012) 

```

<br />

```{r Forests Over Time, echo = FALSE}
#plot forests vs. year
ggplot(line_graph_data, aes(x = year, y = total_forest)) + geom_line(color = "darkgreen") + geom_point() + ggtitle("Woodland Over Time") + xlab("Year") + ylab("Woodland in Thousand Acres") + scale_x_continuous(breaks=c(1945, 1954, 1964, 1974, 1982, 1992, 2002, 2012))
```

```{r Woodland Maps, echo = FALSE}

# 1945 Woodland PCT Map:
wood_map_1945 <- plot_usmap(exclude = c("HI", "AK"), data = wood_1945_state, values = "woodlands_pct", color = 'black') +
  scale_fill_continuous(low = 'white', high = 'brown', name = 'Woodland %', label = scales::comma) +
  theme(legend.position = 'right') +  labs(title = "1945")

# 1964 Woodland PCT Map:
wood_map_1964 <- plot_usmap(exclude = c("HI", "AK"), data = wood_1964_state, values = "woodlands_pct", color = 'black') +
  scale_fill_continuous(low = 'white', high = 'brown', name = 'Woodland %', label = scales::comma) +
  theme(legend.position = 'right') +  labs(title = "1964")

# 1992 Woodland PCT Map:
wood_map_1992 <- plot_usmap(exclude = c("HI", "AK"), data = wood_1992_state, values = "woodlands_pct", color = 'black') +
  scale_fill_continuous(low = 'white', high = 'brown', name = 'Woodland %', label = scales::comma) +
  theme(legend.position = 'right') +  labs(title = "1992")

# 2012 Woodland PCT Map:
wood_map_2012 <- plot_usmap(exclude = c("HI", "AK"), data = wood_2012_state, values = "woodlands_pct", color = 'black') +
  scale_fill_continuous(low = 'white', high = 'brown', name = 'Woodland %', label = scales::comma) +
  theme(legend.position = 'right') +  labs(title = "2012")

grid.arrange(wood_map_1945, wood_map_1964, wood_map_1992, wood_map_2012, ncol=2, nrow=2, top= "U.S. Percentage Woodland 1945-2012")

```

## Special Use and Urban Areas


<br />

```{r Special Areas Over Time, echo = FALSE}
#plot special areas vs. year
ggplot(line_graph_data, aes(x = year, y = total_special)) + geom_line(color = "purple2") + geom_point() + ggtitle("Special Areas Over Time") + xlab("Year") + ylab("Special Area in Thousand Acres") + scale_x_continuous(breaks=c(1945, 1954, 1964, 1974, 1982, 1992, 2002, 2012))
```
 <br />  
```{r Urban Areas Over Time, echo = FALSE}
#plot urban vs. year
ggplot(line_graph_data, aes(x = year, y = total_urban_use)) + geom_line(color = "grey42") + geom_point() + ggtitle("Urban Areas Over Time") + xlab("Year") + ylab("Urban Area in Thousand Acres") + scale_x_continuous(breaks=c(1945, 1954, 1964, 1974, 1982, 1992, 2002, 2012))
```

---

---

---
---


# Employment Analysis and Visualization


```{r Employment Pie Chart Visualizations, echo = FALSE}

attach(PIE_DATA)
# 1959 Employment Pie Chart: 

dt1 <-data.frame(A=c(Ag_workers[1], Other[1], Unemployment[1]),B=c('Ag_workers','Other_workers','Unemployment'))
myLabel = as.vector(dt1$B) 
myLabel = paste(myLabel, "(",dt1$A,' ', round(dt1$A / sum(dt1$A) * 100, 2), "%)", sep = "") 
tu1 <- ggplot(dt1, aes(x = "", y = A, fill = B)) +
  geom_bar(stat = "identity", width = 1) + 
  coord_polar(theta = "y") + 
  labs(x = "", y = "", title = "1959") + 
  theme(axis.ticks = element_blank()) + 
  theme(legend.title = element_blank(), legend.position = "left") + 
  scale_fill_discrete(breaks = dt1$B, labels = myLabel) + 
  theme(axis.text.x = element_blank()) 



# 1964 Employment Pie Chart:

dt2 <-data.frame(A=c(Ag_workers[2], Other[2], Unemployment[2]), B=c('Ag_workers', 'Other_workers', 'Unemployment'))
myLabel = as.vector(dt2$B) 
myLabel = paste(myLabel, "(",dt2$A,' ', round(dt2$A / sum(dt2$A) * 100, 2), "%)", sep = "") 
tu2<- ggplot(dt2, aes(x = "", y = A, fill = B)) +
  geom_bar(stat = "identity", width = 1) + 
  coord_polar(theta = "y") + 
  labs(x = "", y = "", title = "1964") + 
  theme(axis.ticks = element_blank()) + 
  theme(legend.title = element_blank(), legend.position = "left") + 
  scale_fill_discrete(breaks = dt2$B, labels = myLabel) + 
  theme(axis.text.x = element_blank()) 



# 1992 Employment Pie Chart:

dt8 <- data.frame(A=c(Ag_workers[8], Other[8], Unemployment[8]),B=c('Ag_workers','Other_workers','Unemployment'))
myLabel = as.vector(dt8$B) 
myLabel = paste(myLabel, "(",dt8$A,' ', round(dt8$A / sum(dt8$A) * 100, 2), "%)", sep = "") 
tu3<- ggplot(dt8, aes(x = "", y = A, fill = B)) +
  geom_bar(stat = "identity", width = 1) + 
  coord_polar(theta = "y") + 
  labs(x = "", y = "", title = "1992") + 
  theme(axis.ticks = element_blank()) + 
  theme(legend.title = element_blank(), legend.position = "left") + 
  scale_fill_discrete(breaks = dt8$B, labels = myLabel) + 
  theme(axis.text.x = element_blank()) 



# 2012 Employment Pie Chart:

dt12 <-data.frame(A=c(Ag_workers[12], Other[12], Unemployment[12]),B=c('Ag_workers','Other_workers','Unemployment'))
myLabel = as.vector(dt12$B) 
myLabel = paste(myLabel, "(",dt12$A,' ', round(dt12$A / sum(dt12$A) * 100, 2), "%)", sep = "") 
tu4<- ggplot(dt12, aes(x = "", y = A, fill = B)) +
  geom_bar(stat = "identity", width = 1) + 
  coord_polar(theta = "y") + 
  labs(x = "", y = "", title = "2012") + 
  theme(axis.ticks = element_blank()) + 
  theme(legend.title = element_blank(), legend.position = "left") + 
  scale_fill_discrete(breaks = dt12$B, labels = myLabel) + 
  theme(axis.text.x = element_blank()) 

detach(PIE_DATA)

grid.arrange(tu1, tu2, tu3, tu4, ncol=2, nrow=2, top= "U.S. Employment Sector 1959-2012")

```

```{r Employment vs Land Use Visualization, echo= FALSE}



land_labor_graph <- ggplot(labor_land_comb_data, aes(x = Year)) +
  geom_line(aes(y = ag_land_pct), color = "darkred") +
  geom_line(aes(y = ag_emp_pct), color = "steelblue") +
  theme(legend.position = "right") +
  labs(x = "Year", y = "Percentage", 
       title = "U.S. Agricultural Employment and Land Use 1945 - 2012") +
  theme(plot.title = element_text(hjust = 0.5)) 
    
land_labor_graph
```

* Employment has shifted away from the agricultural sector over the last 70 years.






---

---

---
---


# Findings


This employment change happens to coincide with a fall in the land area used for agricultural purposes:
```{r, echo=FALSE}
cropland_pct <- as.integer(combined_data$Total.cropland)/as.integer(combined_data$Total.land)
ag_worker_pct<- as.integer(combined_data$Ag_workers)/as.integer(combined_data$Employment)

cov(cropland_pct,ag_worker_pct)
```
* The co-variance of the percentage of cropland use and percentage of ag_worker is 9.620744e-05,
  which indicates a positive linear relationship between the two variables.

## What happened to the land?
```{r, echo=FALSE}
late_land_use <- all_land_use %>%
  filter(year > 1970)
late_pct_land <- pct_all_land_use %>%
  filter(year > 1970)
```
1. Correlation between croplands and forests is the strongest.
  + If we plow a new field, odds are good we had to clear a few trees to do it.
```{r, echo=FALSE}
cor(pct_all_land_use$cropland_pct, pct_all_land_use$forest_pct)   #correlation between cropland and forests
```

2. Correlation between cropland and urban land also sizable.
  + Suggests urban/suburban sprawl.
  + Interesting: this relationship is weaker than that between farms and forests.
```{r, echo=FALSE}
cor(pct_all_land_use$cropland_pct, pct_all_land_use$urban_pct)    #correlation between cropland and urban land 
```

3. Correlation between urban land and forests, while negative, is very weak.
  + Suggests increases in urban land happen more at the expense of farms than forests.
```{r, echo=FALSE}
cor(pct_all_land_use$urban_pct, pct_all_land_use$forest_pct)      #correlation between urban land and forests```
```

## Did the USDA's changing methodology affect our findings?
```{r, echo=FALSE}
cor(late_pct_land$cropland_pct, late_pct_land$forest_pct)   #correlation between cropland and forests
cor(late_pct_land$urban_pct, late_pct_land$forest_pct)      #correlation between urban land and forests
cor(late_pct_land$cropland_pct, late_pct_land$urban_pct)    #correlation between cropland and urban land 
```
As it turns out, no. If anything, measuring only the later years shows the predicted trends more conclusively.

---
---
---
---

# Conclusions

* Agricultural productivity of land and labor both increased over the last 70 years.
* Land that used to have farms on it has mostly ceded to forests and cities.
* Urbanization leads to some deforestation, but on balance less deforestation than agriculture.
* Future researchers could produce a more rigorous model of land use relationships by examining multiple countries or studying the US by state.
